A Physically Motivated Framework to Compare Merger Timescales\\ of Isolated Low- and High-Mass Galaxy Pairs Across Cosmic Time

### 2024-03-23 paper/main.tex
edited sections 2-4 based on ekta comments, added outline to discussion section

Our sample consists of major (stellar mass ratio $M_{*2}/M_{*1}= 1/4 - 1$) low-mass and high-mass pairs that are isolated but physically associated.
From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of the subhalos of each pair to study their merger timescales.

\subsection{Pair sample}
We begin with an extension of the \paircat{} described in ~\citet{Chamberlain2024}, which consists of a collection of isolated subhalo pairs at each snapshot in TNG100. 
Rather than only considering pairs for $z<4$, we will collect our base sample at each redshift in the simulation (from $z=0-\sim 27$). 
A brief version of the selection routine is transcribed here for completeness, and note that more detail and discussion of the choices for selection criteria are in Sec.~2 of~\citet{Chamberlain2024}.

@@

We require that subhalos constituting a pair meet a minimum subhalo mass criteria of 
\begin{equation*}
    \mbox{\textbf{minimum subhalo mass:}}\,
    \Mhalo > 1\times10^{9}\Msun.
\end{equation*}
at the snapshot of consideration. 
For each halo in the FoF group that passes the minimum subhalo mass criteria, we utilize the \texttt{SUBLINK} catalogs to trace their mass history~\citep{RG2015}, and find the peak halo mass of each subhalo.

Stellar masses are then assigned to each subhalo in the FoF group using the abundance matching prescription of \citet{Moster2013}. 
The peak halo mass and current redshift are used to calculate the stellar mass of each subhalo, so that the abundance matching prescription is of the form $\ms{}=f(\Mpeak,z)$, with $z$ the redshift of consideration.

@@

@@

\subsection{Mergers}
% Plot \#1: Merger fraction either before or after orbits
% Note that it is very interesting that dwarf pairs in the field have high merger fractions, while the likelihood of merger is very low in more dense environments (add citation).

The primary and secondary subhalos of each pair have a `SubfindID' that is used to identify them in the \texttt{Subfind} merger tree catalogs, which contain information about every FoF group and subhalo at each redshift. 
The \texttt{SUBLINK} catalogs track subhalos from one redshift to the next, and thus enable us to track subhalos both backwards and forwards in time from any given redshift.

We utilize the \texttt{DescendantID} field of the \texttt{SUBLINK} catalogs to determine which pairs from our base sample merge before the end of the simulation (at $z=0$) and when the merger occurs. 
The \texttt{DescendantID} field provides the `SubhaloID'\footnote{Note that the `SubhaloID' is distinct from the `SubfindID', and is unique for every subhalo in the merger trees. 
The \texttt{SUBLINK} catalogs provide the associated `SubfindID' of each subhalo.} of the subhalo's descendent in the next (or one of the following) snapshot, if it has one. 
% By following the \texttt{DescendantID}, you follow a subhalo through all of it's mergers to the `root' of the merger tree.

A pair will be classified as a merger if the primary and secondary subhalo share the same \texttt{DescendantID} at the same redshift. 
This means that the primary and secondary subhalo will merge, producing a single descendant subhalo, by the following redshift from the simulation.
If a pair consists of a primary and secondary that never share a \texttt{DescendantID} at a given redshift, the pair is defined as a `non-merger.'

For each merging pair, we define the `merger redshift' as the redshift that immediately follows the first snapshot where the primary and secondary have the same \texttt{DescendantID}. 
For example, if the primary and secondary have different \texttt{DescendantIDs} from $z=6$ to $z=2$, but have the same \texttt{DescendantID} at $z=2$, then the merger must take place some time between $z=2$ and $z=1.9$, which is the next redshift in the simulation. 
We take $z=1.9$ to be the merger redshift.

@@

A given pair from the base sample at redshift $z_n$, which by definition passes all selection criteria at $z_n$, can be followed backwards and forwards in time using the \texttt{SUBLINK} merger trees.  
We track the positions of both the primary and secondary subhalo at each redshift and calculate the physical separations after accounting for the periodic boundary conditions of the simulation box.
In cases where the primary or secondary does not have a defined position at a given redshift, we set the separation at that snapshot to NaN.\footnote{If a subhalo is very small, or is passing through a more massive subhalo, and is unable to reach the density contrast required to be identified as an independent structure by the Subfind algorithm, it will not have a defined position in the Sublink catalogs. The sublink algorithm allows for subhalos to skip a single snapshot, and identifies the `skipped descendent' in the $S_{n+2}$ snapshot, so that the orbit can be evaluated before and after the skip occurs. See Sec.~3 in ~\citet{RG2015} for more details.}

In addition to the physical separation, we also calculate the scaled separation of a pair at each snapshot. We use the definition of scaled separation from~\cite{Chamberlain2024}, where the physical separation is written instead as a fraction, $p$, of the virial radius $\rsep=p\Rvir$, where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the pair's FoF group.\footnote{Where the secondary is not in the same FoF group as the primary, the virial radius used to calculate the scaled separation will remain that of the primary's FoF group, which merging secondaries will eventually re-enter prior to merger.} 
The $\Rvir$ of the pair's FoF group reasonably approximates the virial radius of a halo with a virial mass equal to the combined subhalo mass of the primary and secondary.
This ``scaled separation" is, by construction, a function of mass and redshift, which will account for the mass difference between low-mass and high-mass pairs and for halo growth over time.

@@

\subsubsection{Defining Post-Infall}
While the orbit of a pair may be calculated at very early times, we wish to constrain our orbital analysis to only physically associated pairs, and thus will not consider the orbit of a subhalo pair before they are a part of the same FoF group. Specifically, we will consider only the ``post-infall" part of each pair's orbit.

@@

Figure~\ref{fig:example-orbits} presents a set of example orbits, which shows the wide variety of orbit-types that can be found in the TNG100 simulation for pairs that were originally selected at $z=1.5$. 
The top panel shows the orbits of 5 pairs that merge, where solid lines show the post-infall parts of the orbit, and dashed lines show the pre-infall portions of the orbit, before the secondary and primary ever share a FoF group halo. 
Even amongst galaxies of the same approximate mass and stellar mass ratio, the spread of merger timescales can be very large, with merger timescales between $\sim1-10\Gyr$ from first infall to merger.

Some secondary subhalos can experience first infall and, after a pericentric passage, can return to such distances that they are temporarily assigned a different FoF group halo. 
Because all of our mergers occur when the subhalos are in the same FoF group, the
post-infall orbit definition is more robust for considering the full interaction timescales of merging pairs than considering the orbit only when the FoF group of the primary and secondary are the same. 
% a variety of different orbits

The middle panel of Fig.~\ref{fig:example-orbits} (``Potential Future Mergers") shows the orbits for 3 pairs that are classified as non-mergers, since they do not merge before $z=0$, but which appear likely to merge within a few Gyr past the end of the simulation. 
These orbits can have a variety of orbital periods, and the number of pericentric passages can vary significantly. 
The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past $10\,\Gyr$, while the darker blue orbit has a much shorter period with many close encounters throughout, and 3 close passages in just the past $2\,\Gyr$.

The bottom panel shows the orbits of "Fly-by" interactions (non-mergers) that are unlikely to merge in the near future, if ever.  
Note that we do not split our non-merger category into fly-bys and potential future mergers for any of our following analyses, and such distinctions were made only for the purposes of showing the diversity of orbits that pairs selected at the same snapshot may follow.

\subsubsection{Uniqueness of orbits}
A given pair from the \paircat{} may pass our selection criteria at many redshifts. 
Since we collect the full orbit (the separation at every redshift where the primary and secondary both exist) for each pair and each redshift, the orbits that are collected for pairs that pass our selection criteria at more than one redshift will all result in identical orbits. 
% However, we only want to keep a single instance of any given orbit in our catalog.

To distinguish the collection of unique orbits, each pair is assigned a `pairkey' while constructing their orbit. 
The pairkey is created by concatenating the earliest SubhaloID of the primary and secondary subhalo from the Sublink catalogs, and is unique for each pair of halos. 
After each pair is assigned a unique pairkey, we save the orbit as a function of time only once for a pair that otherwise would have been double/multi-counted in our orbit catalog.

Note that a single subhalo may be a member of many different pairs, but will only have one unique orbit per unique pair.
For example, the primary of a low-mass pair selected at $z=3$ that merges before $z=2$ may be selected via our selection criteria again at $z=1$ with a new secondary companion. 
In this case, both orbits (of the original low-mass pair and the new pair which includes the remnant subhalo from the previous merger) are retained in the orbit catalog, since each orbit is unique.

Removing multiplicity decreases the raw number of orbits collected by collecting an orbit for each pair from the base sample. 
The total number of pairs (including double counting the same pair at multiple redshifts) is 71,429 for low-mass pairs, and 20,824 high-mass pairs. 
However, after removing all redundant orbits, there remain 22,213 low-mass orbits, and 3,029 high-mass orbits.

% For $n_k$ the number of pairs selected at snapshot $S_k$ via the selection criteria/from the base sample, the total number of pairs is given by
% \begin{equation*}
%     \sum_{k=0}^{99}\, n_k
% \end{equation*}
% The total number, and the total number of high-mass pairs is .
% After selecting only unique pairs, there remain

@@

\section{Pair Sample Properties}
\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/6_paircount.png}
    \caption{The number of low-mass (green) and high-mass (pink) unique pairs (or distinct orbits) that are post-infall and pre-merger as a function of redshift. 
    The solid lines show the total number of orbits the are post-infall at a given redshift (including non-mergers and mergers that have yet to merge), while dashed lines show the number of post-infall orbits at a given redshift that will merge before $z=0$.
    We only count the number of orbits ``at" a given redshift if they have not merged, but have experienced first infall at a previous time (i.e. post-infall). 
    The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample via mergers.
    }
    \label{fig:numorbits}
    \end{center}
\end{figure}

@@

\subsection{Number of Orbits}
We calculate the number of orbits (or orbits in progress) by selecting all orbits at a given redshift that are post-infall and either pre-merger or non-mergers.
The number of orbits that result in mergers at a later redshift are selected in the same way, excluding the non-merging orbits.
For example, the total number of orbits at $z=2$ includes the orbits of pairs that passed the selection criteria at \textit{any} redshift, provided they achieved first infall at $z\geq2$ and have a merger redshift $z<2$. 
A single pair (or equivalently, the pair's orbit) orbit may meet the conditions of being a "post-infall + pre-merger" pair for $z=0-3$, and will thus contribute to the number of orbits at all redshifts from $z=0-3$.

Figure~\ref{fig:numorbits} shows the number of low-mass and high-mass orbits as a function of redshift from $z=0-6$. 
Low-mass orbits (green solid line) are most numerous between $z=1.25-2$, while unique high-mass pairs (pink solid line) are most numerous between $2=0-1$.
The dashed lines show the number of unique pairs that merge prior to $z=0$ for each sample. 
The number of pairs that merge decreases to 0 at $z=0$ since 
%pairs selected at $z=0$ by definition have not merged, and thus will not have the chance to, and 
many pairs that exist at low redshift will have merger timescales longer than the remaining time to the end of the simulation.

% removed
% The thin dotted green and pink lines show the number of pairs selected at each snapshot individually, so pairs at $z=1$ that do not pass the snapshot-selection criteria at $z=2$ are not included in the number of pairs at $z=2$, but may still be counted at all snapshots where the selection criteria are met. 
% These lines are similar to the pair count lines for low-mass and high-mass pairs from Fig.~1 in~\cite{Chamberlain2024}, but only include the number of major pairs rather than the total number of pairs.

\subsection{Merger Fraction}
We calculate the merger fraction by dividing the number of merging orbits (dashed lines in Fig.~\ref{fig:numorbits}) by the number of total unique orbits (solid lines) at a given redshift. 
As before, an orbit with first infall at $z=2$ and merger at $z=1$ will be included in the merger fraction calculation for redshifts $z=1-2$.

Figure~\ref{fig:fmerge} shows the fraction of isolated pairs of low-mass (green) and high-mass (pink) pairs that merge before the end of the simulation as a function of redshift. 
At redshifts $z>2$, the merger fraction for low-mass and high-mass pairs is greater than $0.9$.
The merger fraction for both mass ranges declines to $0$ at $z=0$, due to the very low fraction of pairs at low redshift ($z<1$) that have short enough merger timescales to merge before $z=0$. At $z=0$, any remaining pairs are classified as non-mergers since they did not merge \textit{$z=0$}.

@@

% ###########################################################################################
% Results I
% ###########################################################################################

\section{Results: The Mass and Redshift Dependence of Merger Timescales}
We have created catalogs of unique orbits for a large sample of isolated low-mass and high-mass pairs in the TNG100 simulation.
Next, we will calculate the merger timescales, or time until merger, for all of the merging orbits in our catalog. 
In Sec.~\ref{sec:results-timevsep}, we explore how the time until merger changes as a function of the separation of the pair at a variety of redshifts, and study how the time until merger changes between the low-mass sample and high-mass sample. 
In Sec.~\ref{sec:results-timevredshift}, we investigate the median time until merger for all post-infall and pre-merger isolated pairs as a function of redshift. 
We will additionally examine the merger timescale's dependence on a variety of separation criteria.

In the following analysis, we will utilize the orbit catalog, which consists of the unique orbits for all low-mass and high-mass major isolated pairs from the~\citet{Chamberlain2024} study. 
At a given redshift, we will include all merging pairs that are post-infall and pre-merger at that redshift in our analysis. 
However, separation criteria will also be applied to each redshift independently, so a pair that is in the $10-50\,\kpc$ bin at one redshift will not be part of the sample used at redshifts where it's separation is $>50\,\kpc$.

\begin{figure*}[htb]
    \begin{center}
    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevsseplow-2d.png}
    \caption{\kc{Add z=5, z=6 panels. Add colorbar for percentage per panel (and add number of pairs at that redshift) Add low/high label somewhere, is the lower sep set to 10kpc?} The distribution of times until merger for low-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$. 
    There is a positive correlation between separation at the selected redshift and the time until merger. 
    Of orbit-selected merging pairs at $z=1$, the majority are found with separations $\rsep<100\,\kpc$, and will merge within $3-4\, \Gyr$. 
    Merger times are longer for lower redshift pairs, and the distribution of separations is larger.
    The lowest separation pairs have the shortest merger timescales.
    This plot can be used to get estimates for the merger timescale of an isolated low-mass pair at a given redshift. For example, a low-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-6\,\Gyr$, with a most likely time to merger of around $2\Gyr$.
    \kc{add z=5, z=6}
    }
    \label{fig:timevssep-low}
    \end{center}
\end{figure*}

\begin{figure*}[htb]
    \begin{center}
    \includegraphics[width=\textwidth]{plots/bet-on-it/3_Timevssephigh-2d.png}
    \caption{\kc{Add z=5, z=6 panels. Add colorbar. Add low/high label somewhere, check the lower sep set to 10kpc?} The distribution of times until merger for high-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$.
    Similar to the low-mass pairs, there is a positive correlation between the separation of a merging pair at a given redshift and the time until merger, with the lowest separation pairs tending to have the shortest time until merger.
    Of orbit-selected merging pairs at $z=1$, the majority are found with separations $\rsep<200\,\kpc$, and will merge within $0-7\, \Gyr$. 
    This plot can be used to get estimates for the merger timescale of an isolated high-mass pair at a given redshift. 
    For example, a high-mass pair at $z=2$ with $\rsep\sim 150\,\kpc$ will merge in $0.5-4\,\Gyr$, with a most likely time to merger of around $1.5-2\Gyr$.
    }
    \label{fig:timevssep-high}
    \end{center}
\end{figure*}

\subsection{Separation Dependence of Merger Timescales}\label{sec:results-timevsep}
\kc{note: this might go in the appendix as additional info}
% how we calculate for this section
We calculate the time until merger as a function of separation for low-mass and high-mass pairs at a variety of redshifts from $z=0.5-6$. 
At the given redshift, we identify all merging pairs that are post-infall and pre-merger, then find the associated time difference between the given redshift and the redshift of merger for each pair.

% low mass example
For example, at $z=1.5$, there are 5,787 unique orbits that are post-infall and will merge prior to $z=0$. Of those, and 2,458 (42.47\%) have low separations between $\rsep=10-50\,\kpc$, and 2,207 of those low-separation pairs will merge in the next $2\,\Gyr$. 
A compilation of this data is available in Table~\ref{tab:timevssepvsz}, where we list the fraction of post-infall and pre-merger pairs within 3 separation bins $(10-50, 10-70, 10-100)\,\kpc$, and the fraction within each of those separation bins that merge in 4 different timescale bins $(0-1, 1-2, 2-3, 3+)\,\Gyr$, for redshifts of $z=(0.5,1,1.5,2,3,4,5,6)$..

>>>>

Our sample consists of major (stellar mass ratio $M_{*2}/M_{*1}= 1/4 - 1$) low-mass and high-mass pairs that are isolated but physically associated as in \citet{Chamberlain2024}.
From this sample, we will determine the fraction of pairs at each snapshot that merge before $z=0$, and track the orbits of each pair to study their merger timescales.

\subsection{Pair sample}
We begin with an extension of the \paircat{} described in ~\citet{Chamberlain2024}, which consists of a collection of isolated subhalo pairs at each snapshot in TNG100. 
We collect our base sample at each redshift in the simulation (from $z=0-\sim 27$). 
A brief version of the selection routine is transcribed here for completeness, and we refer readers to Sec.~2 of~\citet{Chamberlain2024} for more detail and a discussion of the selection criteria choices.

@@

We require that subhalos constituting a pair meet a minimum subhalo mass criteria of 
\begin{equation*}
    \mbox{\textbf{minimum subhalo mass:}}\,
    \Mhalo > 1\times10^{9}\Msun.
\end{equation*}
at the snapshot of consideration. 
For TNG100, this ensures that subhalos are resolved into over 130 particles which should be robustly identified in the SUBFIND and SUBLINK catalogs.
For each halo in the FoF group that passes the minimum subhalo mass criteria, we utilize the \texttt{SUBLINK} catalogs to find the peak halo mass of each subhalo \citep{RG2015}.

Stellar masses are assigned to each subhalo in the FoF group using the abundance matching prescription of \citet{Moster2013}. 
The peak halo mass and current redshift are used to calculate the stellar mass of each subhalo via the abundance matching prescription  $\ms{}=f(\Mpeak,z)$, where $z$ is the redshift of consideration.

@@

A subhalo can belong to different pairs at different redshifts. 
For example, the primary of a pair that merges at $z=2$ can be selected with a different secondary at $z=1$, constituting a new pair. 
More detail regarding the uniqueness of pairs and orbits is discussed in~\ref{sec:methods-unique}.

@@

\subsection{Mergers}
% Plot \#1: Merger fraction either before or after orbits
% Note that it is very interesting that dwarf pairs in the field have high merger fractions, while the likelihood of merger is very low in more dense environments (add citation).
\label{subsec:mergers}
The primary and secondary subhalos of each pair have a `SubfindID' that is used to identify them in the \texttt{Subfind} merger tree catalogs, which contain information about every FoF group and subhalo at each redshift. 
The \texttt{SUBLINK} catalogs track subhalos from one redshift to the next, and thus enable us to track subhalos both backwards and forwards in time from any given redshift.

We utilize the \texttt{DescendantID} field of the \texttt{SUBLINK} catalogs to determine which pairs from our base sample merge before the end of the simulation (at $z=0$) and when the merger occurs. 
The \texttt{DescendantID} field provides the `SubhaloID'\footnote{Note that the `SubhaloID' is distinct from the `SubfindID', and is unique for every subhalo in the merger trees. 
The \texttt{SUBLINK} catalogs provide the associated `SubfindID' of each subhalo.} of the subhalo's descendent in the next (or one of the following) snapshots, if it has one. 
% By following the \texttt{DescendantID}, you follow a subhalo through all of it's mergers to the `root' of the merger tree.

In this analysis, a pair is classified as a merger if the primary and secondary subhalo share the same \texttt{DescendantID} at the same redshift. 
This means that the primary and secondary subhalo have merged such that SUBFIND can no longer distinguish them as two separate halos, therefore yielding a single descendant subhalo.
If a pair consists of a primary and secondary that never share a \texttt{DescendantID} at a given redshift, the pair is defined as a `non-merger.'

For each merging pair, we define the `merger redshift' as the redshift that immediately follows the first snapshot where the primary and secondary have the same \texttt{DescendantID}. 
For example, if the primary and secondary have different \texttt{DescendantIDs} from $z=6$ to $z=2$, but have the same \texttt{DescendantID} at $z=2$, then the merger must take place some time between $z=2$ and $z=1.9$, which is the next redshift corresponding to a snapshot in the simulation. 
In this example, we take $z=1.9$ to be the merger redshift.

@@

A given pair from the base sample at redshift $z_n$, which by definition passes all selection criteria at $z_n$, can be followed backwards and forwards in time using the \texttt{SUBLINK} merger trees.  
We track the positions of both the primary and secondary subhalo at each redshift and calculate the physical separations after accounting for the periodic boundary conditions of the simulation box.
In cases where the primary or secondary does not have a defined position at a given redshift, we do not compute a value for the separation at that redshift.\footnote{If a subhalo is very small, or is passing through a more massive subhalo, and is unable to reach the density contrast required to be identified as an independent structure by the Subfind algorithm, it will not have a defined position in the Sublink catalogs. The sublink algorithm allows for subhalos to skip a single snapshot, and identifies the `skipped descendent' in the $S_{n+2}$ snapshot, so that the orbit can be evaluated before and after the skip occurs. See Sec.~3 in ~\citet{RG2015} for more details.}

In addition to the physical separation, we also calculate the scaled separation of a pair at each redshift. We use the definition of scaled separation from~\cite{Chamberlain2024}, where the physical separation is written instead as a fraction, $p$, of the virial radius $\rsep=p\Rvir$, where $\rsep$ is the physical separation in $\kpc$, and $\Rvir$ is the virial radius of the pair's FoF group.\footnote{Where the secondary is not in the same FoF group as the primary, the virial radius used to calculate the scaled separation will remain that of the primary's FoF group, which merging secondaries will eventually re-enter prior to merger.} 
The $\Rvir$ of the pair's FoF group reasonably approximates the virial radius of a halo with a virial mass equal to the combined subhalo mass of the primary and secondary.
This ``scaled separation" is, by construction, a function of mass and redshift, which will account for the mass difference between low-mass and high-mass pairs and for halo growth over time.

@@

\subsubsection{Defining Post-Infall}
While the orbit of a pair may be calculated at very early times, we wish to constrain our orbital analysis to only physically associated pairs, and thus will not consider the orbit of a subhalo pair before they belong to the same FoF group. Specifically, we will consider only the ``post-infall" part of each pair's orbit.

@@

Figure~\ref{fig:example-orbits} presents a set of example orbits, which shows the wide variety of orbit-types that can be found in the TNG100 simulation for pairs that were originally selected at $z=1.5$. 
The top panel shows the orbits of five pairs that merge, where solid lines show the post-infall parts of the orbit, and dashed lines show the pre-infall portions of the orbit, before the secondary and primary ever share a FoF group halo. 
Even amongst galaxies of the same approximate mass and stellar mass ratio, the spread of merger timescales can be very large, with merger timescales between $\sim1-10\,\Gyr$ from first infall to merger,  indicated by the triangles.

For some pairs, the secondary subhalos can experience first infall and, after a pericentric passage, can return to such distances that they are temporarily assigned a different FoF group halo than their primary. 
Because all of our mergers occur when the subhalos are in the same FoF group by definition, in these cases, we use the full orbit starting when the subhalos are first identified as members of the same FoF group through when they merge, as described in Section \ref{subsec:mergers}.
This definition is more robust for considering the full interaction timescales of merging pairs.
%the post-infall orbit definition is more robust for considering the full interaction timescales of merging pairs than considering the orbit only when the FoF group of the primary and secondary are the same. }
% a variety of different orbits

The middle panel of Fig.~\ref{fig:example-orbits} (``Potential Future Mergers") shows the orbits for three pairs that are classified as non-mergers, since they do not merge before $z=0$, but which appear likely to merge within a few Gyr past the end of the simulation. 
These orbits can have a variety of orbital periods, and the number of pericentric passages can vary significantly. 
The two lighter blue orbits are very long period orbits with 1-2 pericenter passages in the past $10\,\Gyr$, while the darker blue orbit has a much shorter period with many close encounters throughout, and three close passages in just the past $2\,\Gyr$.

The bottom panel shows the orbits of ``Fly-by" interactions (non-mergers) that are unlikely to merge in the near future, if ever.  
Note that we do not split our non-merger category into fly-bys and potential future mergers for any of our following analyses, and such distinctions were made only for the purposes of showing the diversity of orbits that pairs selected at the same snapshot may follow.

\subsubsection{Uniqueness of orbits}
\label{sec:methods-unique}
Since we collect the orbit (the separation at every redshift where the primary and secondary both exist) for each pair at all redshifts, a singular orbit will be selected as many times as the number of redshifts (or snapshots) where that pair exists. However, we only want to keep a single instance of any given orbit in our catalog so we can track the merger timescale of the pair.

To distinguish the collection of unique orbits, each pair is assigned a `pairkey' while constructing their orbit. 
The pairkey is created by concatenating the earliest SubhaloID of the primary and secondary subhalo from the Sublink catalogs, and is unique for each pair of halos. 
After each pair is assigned a unique pairkey, we only keep one instance of an orbit per pair to avoid double/multi-counting in our orbit catalog.

Note that a single subhalo may be a member of many different pairs, but will only have one unique orbit per unique pair.
For example, the primary of a low-mass pair selected at $z=3$ that merges before $z=2$ may be selected via our selection criteria again at $z=1$ with a new secondary companion. 
In this case, both orbits (of the original low-mass pair and the new pair which includes the primary subhalo from the previous merger) are retained in the orbit catalog, since each orbit is unique.

The total number of pairs (including double counting the same pair at multiple redshifts) is 71,429 for low-mass pairs, and 20,824 high-mass pairs. 
However, after removing all redundant orbits, there remain 22,213 low-mass orbits, and 3,029 high-mass orbits that each correspond to a unique pair of galaxies.

@@

\section{Pair Sample Properties}

\begin{figure}[htb]
    \begin{center}
    \includegraphics[width=\columnwidth]{plots/bet-on-it/6_paircount.png}
    \caption{The number of low-mass (green) and high-mass (pink) unique pairs (or distinct orbits) that are post-infall and pre-merger as a function of redshift. 
    The solid lines show the total number of pairs the are post-infall at a given redshift (including non-mergers and mergers that have yet to merge), while dashed lines show the number of post-infall orbits at a given redshift that will merge before $z=0$.
    We only count the number of orbits ``at" a given redshift if they have not merged, but have experienced first infall at a previous time (i.e. post-infall). 
    The decrease in the number of unique low-mass pairs from $z=2$ to $z=0$ is due to pairs being removed from the sample via mergers.\kc{change ylabel to "pairs"}
    }
    \label{fig:numorbits}
    \end{center}
\end{figure}

@@

\subsection{Number of Pairs}
We calculate the number of pairs (i.e. orbits in progress) by selecting all pairs at a given redshift that are post-infall and have not merged.
% The number of orbits that result in mergers at a later redshift are selected in the same way, excluding the non-merging orbits.
%For example, the total number of orbits at $z=2$ includes the orbits of pairs that passed the selection criteria at \textit{any} redshift, provided they achieved first infall at $z\geq2$ and have a merger redshift $z<2$. 
A single pair (or equivalently, the pair's orbit) may meet the conditions of being a "post-infall + pre-merger" pair for $z=0-3$, and will thus contribute to the number of pairs at all redshifts from $z=0-3$.

Figure~\ref{fig:numorbits} shows the number of low-mass and high-mass pairs as a function of redshift from $z=0-6$. 
Low-mass pairs (green solid line) are most numerous between $z=1.25-2$, while unique high-mass pairs (pink solid line) are most numerous between $z=0-1$.
The dashed lines show the number of unique pairs that merge prior to $z=0$ for each sample. 
The number of pairs that merge decreases to 0 at $z=0$ since many pairs that exist at low redshift will have merger timescales that span beyond the end of the simulation (i.e., $z=0$).

When a pair experiences first infall and thus begins to contribute to the number of pairs at later redshifts, we refer to such as pairs being ``added to the sample".
Likewise, mergers at a given redshift will decrease the number of pairs counted at that redshift, since the pair is then post-merger and thus not included in the count.

% The difference between high and low mass
The number of low-mass pairs continues to increase from $z=6$ until around $z=2$, at which point the number of added pairs at a given redshift slows. 
The decrease in the number of pairs after $z=2$ means that the number of low-mass pairs that are merging at each redshift is larger than the rate at which pairs are being added to the sample.

The number of high-mass pairs continues increasing until $z=1$, at which point it remains approximately constant from $z=0-1$ at $\sim1000$ pairs. 
There is a slight decrease in the number of high mass pairs at the very lowest redshifts $z<0.1$, where the mergers begin to outnumber the added pairs at each redshift.

\subsection{Merger Fraction}
We calculate the merger fraction by dividing the number of pairs that merge (dashed lines in Fig.~\ref{fig:numorbits}) by the number of total unique orbits (solid lines) at a given redshift, including all pairs that do and do not merge. 
As before, an orbit with first infall at $z=2$ and merger at $z=1$ will be included in the merger fraction calculation for redshifts $z=1-2$.

Figure~\ref{fig:fmerge} shows the fraction of isolated pairs of low-mass (green) and high-mass (pink) pairs that merge before the end of the simulation as a function of redshift. 
At redshifts $z>2$, the merger fraction for low-mass and high-mass pairs is greater than $0.9$.
The merger fraction for both mass ranges declines to $0$ at $z=0$, due to the very low fraction of pairs at low redshift ($z<1$) that have short enough merger timescales to merge before $z=0$.

% difference/similiarity between low and high
The merger fraction of low-mass and high-mass pairs is the same from $z=0-0.5$ and $z>2.5$. 
Between $z=0.5-2.5$, the high-mass merger fraction is higher than the low-mass merger fraction, and the knee of the merger fraction occurs at a lower redshift.
This is likely because the number of low-mass pairs continues to increase before turning over and beginning to decrease at $z\sim2$.
When the number of pairs stops increasing and beings to decrease, more pairs that will not have time to merge before $z=0$ are being added, and many merging pairs are being removed from the sample, thus decreasing the merger fraction. 
Since the turnover for the number of high-mass pairs occurs at a lower redshift, the knee of the merger fraction is likewise shifted to lower redshifts.

@@

% ###########################################################################################
% Results I
% ###########################################################################################
\begin{figure*}[htb]
    \begin{center}
    \includegraphics[width=0.7\textwidth]{plots/bet-on-it/3_Timevsseplow-2d.png}
    \caption{\kc{Add z=5, z=6 panels. Add colorbar for percentage per panel (and add number of pairs at that redshift) Add low/high label somewhere, is the lower sep set to 10kpc?} The distribution of times until merger for low-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$. 
    There is a positive correlation between separation at the selected redshift and the time until merger. 
    Of orbit-selected merging pairs at $z=1$, the majority are found with separations $\rsep<100\,\kpc$, and will merge within $3-4\, \Gyr$. 
    Merger times are longer for lower redshift pairs, and the distribution of separations is larger.
    The lowest separation pairs have the shortest merger timescales.
    This plot can be used to get estimates for the merger timescale of an isolated low-mass pair at a given redshift. For example, a low-mass pair at $z=2$ with $\rsep\sim 75\,\kpc$ will merge in $0-6\,\Gyr$, with a most likely time to merger of around $2\Gyr$.
    \kc{add z=5, z=6, add horizontal line for the time remaining in the sim in panel 1}
    }
    \label{fig:timevssep-low}
    \end{center}
\end{figure*}

\begin{figure*}[htb]
    \begin{center}
    \includegraphics[width=0.7\textwidth]{plots/bet-on-it/3_Timevssephigh-2d.png}
    \caption{\kc{Add z=5, z=6 panels. Add colorbar. Add low/high label somewhere, check the lower sep set to 10kpc?} The distribution of times until merger for high-mass pairs as a function of physical separation at $z=0.5,1,1.5,2,3,\mbox{and }4$.
    Similar to the low-mass pairs, there is a positive correlation between the separation of a merging pair at a given redshift and the time until merger, with the lowest separation pairs tending to have the shortest time until merger.
    Of orbit-selected merging pairs at $z=1$, the majority are found with separations $\rsep<200\,\kpc$, and will merge within $0-7\, \Gyr$. 
    This plot can be used to get estimates for the merger timescale of an isolated high-mass pair at a given redshift. 
    For example, a high-mass pair at $z=2$ with $\rsep\sim 150\,\kpc$ will merge in $0.5-4\,\Gyr$, with a most likely time to merger of around $1.5-2\Gyr$.
    }
    \label{fig:timevssep-high}
    \end{center}
\end{figure*}

\section{Results: The Mass and Redshift Dependence of Merger Timescales}
Using our sample of isolated low-mass and high-mass pairs in the TNG100 simulation, we calculate the merger timescales, or time until merger, for all of the merging pairs in our catalog. 
In Sec.~\ref{sec:results-timevsep}, we explore how the merger timescale changes as a function of the separation of the pair across redshifts, and how the merger timescales differ between the low-mass and high-mass pairs. 
In Sec.~\ref{sec:results-timevredshift}, we investigate the median time until merger for all post-infall and pre-merger isolated pairs as a function of redshift. 
We will additionally examine the merger timescale's dependence on a variety of separation criteria. 
Separation criteria are applied to pairs at each redshift independently, such that a pair in the $10-50\,\kpc$ bin at one redshift will not be part of the sample used at redshifts where its separation is $>50\,\kpc$.

\subsection{Separation Dependence of Merger Timescales}\label{sec:results-timevsep}
% how we calculate for this section
We calculate the time until merger as a function of separation for low-mass and high-mass pairs at a variety of redshifts from $z=0.5-6$. 
Pairs with small separations may be undergoing a pericentric passage, or very close to merging if on a radial orbit.
Binning the pairs by separation, we can study how the merger timescale changes for pairs at different points in their orbits.

% low mass example
We bin our pair sample by merger timescale, in bins of $0.5\Gyr$, and separation in bins of width $10\,\kpc$.
Figures~\ref{fig:timevssep-low} and~\ref{fig:timevssep-high} show the 2D distributions of time until merger vs. separation for low-mass and high-mass pairs at redshifts $z=(0.5,1,1.5,2,3,4,5,6)$. 
The horizontal line in the first panel shows the time remaining in the simulation until $z=0$, above which no merging pairs exist.

Fig.~\ref{fig:timevssep-low} presents a heatmap distribution of the density of low-mass pairs in merger timescale-separation space. 
We find that the time until merger for low-mass pairs is positively correlated with the separation of the pair at each redshift. 
However, the slope of the correlation decreases at higher redshifts.
Pairs at $z=4$ and separations between $50-100\,\kpc$ tend to merge in less time than pairs with the same separations at $z=0.5$.

In addition, the spread of times until merger and separations is smaller at higher redshift. 
The spread of merger timescales at $z=1$ goes from $0-8\,\Gyr$, while pairs at $z=4$ have merger timescales between $0-3\,\Gyr$, and likewise for the separations spanned by pairs at both redshifts. 
This is due to the growth of halo size over time, since more pairs will be post-infall at lower redshifts where the virial radius of the FoF groups is larger.

Likewise, Fig.~\ref{fig:timevssep-high} presents the distribution for high-mass subhalos.
The trends between merger timescale and separation are roughly the same as low-mass pairs, with a positive correlation between the two parameters that decreases in slope from low redshift to high redshift. 
The high-mass pairs also have a larger spread in the distribution of merger timescales and separations at low redshift than at higher redshift.
This is once again caused by the change in the FoF halo size over time.
